---
title: "Code for HW 5"
author: Maddie Manzagol 
date: June 6, 2023 
format: 
  html:
    toc: true
    toc-location: left
    code-fold: true
    theme: yeti 
execute:
  message: false
  warning: false
editor_options: 
  markdown: 
    wrap: 72
---
# Link to Github

[github](https://github.com/maddie-manzagol/ENVS-193DS_homework-05.git)

# Introduction

Sarracenia, commonly known as pitcher plants, are carnivorous wetland
flora native to North America that thrive in sunny, wet, nutrient-poor
habitats (Horner). Sarracenia may be of interest to people because their
predation strategies can have indirect as well as direct impacts on the
population dynamics and community structure of freshwater plankton and
bacterial communities due to their interactions at higher trophic levels
which control bacterial densities and community structure
(Cochran-Stafira). Sarracenia may also be of interest to people because
their carnivorous behavior causes them to not follow the 'world-wide
leaf economics spectrum' which is defined as an allometric relationship
between nutrients, maximal net photosynthetic rates, and physiology
(notably leaf size) (Farnsworth and Ellison). The reason their leaf
traits diverge from the leaf economics spectrum is due to the fact that
carnivorous plants derive most of their nutrients from capturing and
digesting prey, implying low photosynthetic rates and low photosynthetic
nitrogen and phosphorous use efficiencies (Farnsworth and Ellison).
Since prey capture is their main source of nutrient uptake, prey capture
is known to be indicative of their fitness in the sense that increased
prey capture aids in survival, growth, and reproduction (Horner). Due to
the impact on the individual fitness of Sarracenia plants, it is
important to understand how Sarracenia traits aid in prey capture as
well as how Sarracenia react to changes in the abundance of prey
(Horner). Additionally, it is valuable to predict individual biomass of
Sarracenia based on morphological, physiological, or taxonomic
characteristics because the prediction models offer insight into their
growth patterns, resource allocation, and ecological roles, potentially
shedding light on larger ecological processes, as well as, how humans
can adopt effective management practices. This study investigates the
effects of prey availability on photosynthetic rate (Amass), chlorophyll
fluorescence, growth, architecture, and foliar nutrient and chlorophyll
content of ten pitcher plant (Sarracenia) species. The study questions
how prey availability impacts Sarracenia growth and health, as well as,
if morphological and physiological allometric relationships of
Sarracenia resemble allometric relationships of non-carnivorous plants
when nutrients are not a limiting resource. These questions formulated
the hypothesis that increased prey availability would stimulate Amass,
increase photosynthetic N- and P-use efficiencies, increase relative
biomass allocation to photosynthetically efficient, non-predatory
phyllodes rather than pitchers, and thereby increase overall plant
biomass and reduce stress-related chlorophyll fluorescence.

# Methods

a.  Sampling Methods

120 Sarracenia plants ranging from small, intermediate, and large
species size were used for the study were collected from the sampling
site: Harvard Forest Greenhouse (Ellison and Farnsworth). To conduct the
study, plants were fed once a week for 7 weeks straight (Ellison and
Farnsworth). Two plants of each Sarracenia species were assigned to one
of six feeding levels in a regression design ranging from 0 - 0.25g of
finely ground wasps per feeding (for small species), 0 - 0.5g (for
species of intermediate sizes), and 0 - 1.0g (for large species)
(Ellison and Farnsworth). "Aboveground size" and "Amass" were measured
prior to commencing treatments, meanwhile all other parameters were
measured during the duration of the study (Ellison and Farnsworth).

b.  Description of data organization and processing

In order to organize the data after reading in the necessary packages,
it is important to see what data was measured. To do so, the full data
set was read in using the read_csv and the here function, then to make
the names of the objects within the data set more clear and too only
include the objects of interest the clean_names and select functions
were applied. The data set now appears readable with the data set only
reporting values observed for totmass, species, feedlevel, sla,
chlorophyll, amass, num_lvs, num_phylls

Read in packages

```{r libraries}
# should haves (from last week)
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
# would be nice to have
library(corrplot)
library(AICcmodavg)
library(GGally)
```

Read in the csv file

```{r read-in-data}
#read in the data
plant <- read_csv(here("data","knb-lter-hfr.109.18 (1)", "hf109-01-sarracenia.csv")) %>% 
  #make the column names clear
  clean_names() %>% 
  #selecting the columns of interest
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)
```

c.  Visualization of missing observations with an accompanying caption
    and discussion

Discussion: Given that there are many missing data throughout multiple
parameters, it is important to identify where the missing data is so
that one can identify where bias potentially is within statistical
analysis as it might interfere with results.

```{r missing-data-visualization}
#visualize the missing data
gg_miss_var(plant)
```

Caption: The gg_miss_var function highlights missing data within the
"plant" data set. The plot highlights missing data in the chlorophyll,
amass, sla, num_phylls, and num_lvs columns meaning that the only
parameters without missing data includes totmass, species, and
feedlevel.

Subsetting the data by dropping NAs:

```{r subset-drop-NA}
#new data set with dropped NA values
plant_subset <- plant %>% 
  drop_na(sla, chlorophyll, amass, num_lvs, num_phylls)
```

Create a correlation plot:

d.  Visualization of Pearson's correlation between variables with an
    accompanying caption and discussion

Discussion: After calculating pearson's r for the different variables
within the plant_subset data frame, it was determined that none of the
variables have a super strong linear association as the highest
correlation value (r) was between amass and sla (0.32). The correlations
between sla and feedlevel (-0.3), chlorophyll and sla (-0.22), num_lvs
and feedlevel (-0.07), num_phylls and sla (-0.3), and num_phylls and
chlorophyll (-0.24) are all negatively correlated; meanwhile, the
correlations between chlorophyll and feedlevel (0.16), amass and
feedlevel (0.13), num_phylls and feedlevel (0.25), sla and amass (0.32),
and sla and num_lvs (0.09) are all postively correlated. It is also
worthy to note that there was no correlation between amass and
num_phylls (0).

```{r correlation-plot}
#calculate pearson's r only for numerical values only
plant_cor <- plant_subset %>% 
  select(feedlevel:num_phylls) %>% 
  cor(method = "pearson")

#creating a correlation plot
corrplot(plant_cor,
         #change the shape of what's in the cells
         #point to the left=negative relationship, right=positive relationship
         method = "ellipse",
         #adds the correlation values in black over the elipses
         addCoef.col= "black")
```

Caption: To determine the relationship between numerical variables in
our dataset, we calculcated pearson's r and visually represented the
correlation using a correlation plot. We compared the variables
feedlevel, sla, chlorophyll, amass, num_lvs, and num_phylls.

e.  Visualization of relationships between variables (not correlation)
    with an accompanying caption and discussion

Discussion: After visualizing the relationships between each pair of
variables, the ggpair function reveals the different relationships
between different varibales depending on the type of data each variable
is. Box plots are used to represent relationships between species,
categorical data, and the rest of the variables, continuous or discrete.
The scatter plots and line graphs highlight relationships between
continuous and discrete data.

Create a plot of each variable compared against the others

```{r pair-plot}
#visualize the relationships between variables, not correlation 
plant_subset %>% 
  select(species:num_phylls) %>% 
  ggpairs()
```

Caption: In order to visualize the relationship between the variables
the ggpairs function was applied. The ggpairs function builds a
scatterplot matrix that displays scatterplots of each pair of numeric
values, while also displaying pearson's r on the right and the variable
distribution on the diagonal.

Starting Regression here:

f\. Creating the null and full model with accompanying discussion

To determine how species and physiological characteristics perditc
biomass, we fit multiple linear models. The null model creates a
regression model that assumes no realtionship between totmass and the
predictor variables. On the other hand, the full model creates a
regression model that assumes a relationship between totmass and the
predictor variables in order to predict biomass.

```{r null-and-full-models}
#create linear regression models for null and full model
null <- lm(totmass ~ 1, data= plant_subset)
full <- lm(totmass ~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data= plant_subset)
```

g.  Visual and statistical assumption checks for the full model with
    accompanying discussion

We visually assesed normaility and homoskedascitity using diagnostic
plots for the full model and found that the data is appears normal, but
not homoskedastic because the data is not scattered entirely random. We
also tested for normality using the Shaprio-Wilk test and found
non-normality of residuals. We then retested for homoscedasticity using
the Breusch-Pagan test and detected non-constant error variance meaning
the data is heteroscedastic.

```{r full-diagnoistics}
#check diagnostics
par(mfrow = c(2,2))
plot(full)
```

```{r}
#check assumptions
check_normality(full)
check_heteroscedasticity(full)
```

h.  Explanation of any transformation and justification

After visually assesing normaility and homoscedascitity using diagnostic
plots for the full model, we found that the data appeared to be
non-normal and heteroscedasitic. In order to reduce the skewness of
variable of interest and achieve normality, we applied a log
transformation. Once we did this we were able to manipulate the
residuals in a manner that reported normal distribution and
homoscedastic variance.

```{r log transformation}
#create linear regression models for log transformation of null and full model
null_log <- lm(log(totmass)~1, data = plant_subset)
full_log <- lm(log(totmass)~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data= plant_subset)

#plot and check assumptions 
plot(full_log)
check_normality(full_log)
check_heteroscedasticity(full_log)
```

i.  Model construction with visual and statistical assumption checks for
    three additional models with justification

To address the question, what set of predictor variables best explains
the response, we applied three more linear models using the predictor
variables species, feedlevel, and num_lvs. The species variable was
chosen as a predictor variable because it is probable that different
species have different biomasses. The model comparing species as
predictor variable and totmass as the response variable was found to be
normal and homoscedastic. The feedlevel variable was chosen as a
predictor variable because different feed levels are associated with
different nutrient level. Different nutrient levels are indicative of
growth and overall health which contributes to individual's biomass. The
num_lvs variable was chosen as a predictor variable because the more
leaves an individual has, the more biomass an individual has. The models
comparing feedlevel and num_lvs to predict totmass, respectively, both
were found to be non-normal, yet homoscedastic.

Model with predictor variable: species

```{r}
#New linear regression model with species are predictor
model2_log <- lm(log(totmass)~ species, data= plant_subset)
```

check assumptions for model 2:

```{r assumptions}
#plot the model and check for normality and heteroscedasticity
plot(model2_log)

check_normality(model2_log)
check_heteroscedasticity(model2_log)
```

Model with predictor variable: feedlevel

```{r}
#New linear regression model with feedlevel are predictor
model3_log <- lm(log(totmass)~ feedlevel, data = plant_subset)
```

check assumptions for model 3:

```{r}
#plot the model and check for normality and heteroscedasticity
plot(model3_log)

check_normality(model3_log)
check_heteroscedasticity(model3_log)
```

Model with predictor variable: num_lvs

```{r}
#New linear regression model with num_lvs are predictor
model4_log <- lm(log(totmass)~ num_lvs, data = plant_subset)
```

check assumptions for model 4:

```{r}
#plot the model and check for normality and heteroscedasticity
plot(model4_log)

check_normality(model4_log)
check_heteroscedasticity(model4_log)
```

j.  Variance inflation factor check for the full model with accompanying
    discussion 

We evaluated muticollinarity by calculating the generalized variance
inflation factor and determined that this model didn't display any
multicollinearity because no values were larger than 5.

```{r calculate-vif}
#from the car package
car::vif(full_log)
```

k.  Model comparison using the null model, full model, and additional
    models from part 7g

compare models using Akaike's Information criterion (AIC) values:

```{r AIC-value}
#generate AIC values
AICc(full_log)
AICc(model2_log)
AICc(model3_log)
AICc(model4_log)
AICc(null_log)

#rank AIC values
MuMIn::AICc(full_log, model2_log, model3_log, model4_log, null_log)
MuMIn::model.sel(full_log, model2_log, model3_log, model4_log, null_log)
```

We compared models using AIC and chose the model with the lowest value,
which was full_log model.

# Results

We found that the full_log model including species, feedlevel, sla,
chlorophyll, amass, num_lvs, and num_phylls predictors best predicted
totmass because the full_log model yields the lowest AIC value
indicating that this model has the best predictive capacity while
remaining as simple as possible. Overall, the full_log model was
statistically significant (R\^2= 0.8687, F(15, 87)= 38.38, p\<0.001).
With all else held constant, the full_log model indicates a significant
relationship between totmass and num_lvs (p\< 0.001). Comparing to the
alabamensis species of pitcher plants, the leucophylla species are
likely to have a biomass that is 1.7 higher (95% confidence level,
\[1.25, 2.16\]) and the psittacina species are likely to have a biomass
that is 1.6 lower (95% confidence level, \[-2.05,-1.23\]). However it
was found that some of the predictors such as, the jonesii and purpurea
species, amass, and num_phylls, did not significantly predict totmass
(p= 0.107, p=0.155, p= 0.436, p= 0.446, respectively). In conclusion,
our model highlights that variables such as number of leaves, feed
level, specific leaf area, chlorophyll, and specific species type (lata,
flava, leucophylla, minor, psittacina, rosea, and rubra) are all
significant variables in predicting the biomass of one individual pitch
plant. Focusing on the number of leaves, individuals producing more
pitchers and phyllodes will predictably accumulate more biomass than
individuals who produce less pitches and phyllodes.

```{r}
#create a summary table with summary statistics
summary(full_log)
tidy(full_log, conf.int = TRUE, conf.level = 0.95)
```

```{r}
#use ggpredict() to back transform estimates and save it as an object called "model_pred"
model_pred <- ggpredict(full_log, terms = "num_lvs", back.transform = TRUE)

#plot "model_pred" object 
plot(model_pred, add.data = TRUE)

```

Caption: We used the ggpredicts function to predict biomass from the
predictor variable num_lvs. We did by computing back transformating the
predictor variable num_lvs to create an estimate for biomass. We found a
postive correlation between the predictor variable, num_lvs, and the
response variable, totmass, indicating that as the number of leaves
increases, so does biomass.

```{r}
table <- tidy(full_log, conf.int = TRUE) %>% 
    #using mutate to change the pvalue numbers if they are really small
  #using mutate to chnage the estimates, standard error, and t-statistics to round to 4 digits
  mutate(across(estimate:conf.high,~round(.x,digits=4))) %>% 
  mutate(statistic=round(statistic, digits = 1)) %>% 
  mutate(p.value=case_when(p.value<0.001~"<0.001",.default = as.character(p.value))) %>%
  #fix this
  #make it into a flex table 
  flextable() %>% 
  set_header_labels(term="Term",
                    estimate="Estimate",
                    std.error="Standard Error",
                    statistic="Statistic",
                    p.value="p-value",
                    conf.low="Lower Confidence Level",
                    conf.high="High Confidence Level") %>% 
  # fit it to the viewer 
  autofit()
  

table
```

# Bibliography

Cochran-Stafira, D.L. and von Ende, C.N. (1998), INTEGRATING BACTERIA
INTO FOOD WEBS: STUDIES WITH SARRACENIA PURPUREA INQUILINES. Ecology,
79: 880-898.
[doi](https://doi.org/10.1890/0012-9658(1998)079\[0880:IBIFWS\]2.0.CO;2)
(Accessed 2023-06-04).

Ellison, A. and E. Farnsworth. 2021. Effects of Prey Availability on
Sarracenia Physiology at Harvard Forest 2005 ver 18. Environmental Data
Initiative.
[doi](https://doi.org/10.6073/pasta/26b22d09279e62fd729ffc35f9ef0174) (Accessed
2023-06-04).

Farnsworth, Elizabeth J., and Aaron M. Ellison. "Prey Availability
Directly Affects Physiology, Growth, Nutrient Allocation and Scaling
Relationships among Leaf Traits in 10 Carnivorous Plant Species."
Journal of Ecology, Vol. 0, No. 0, 2007,
[doi](https://doi.org/10.1111/j.1365-2745.2007.01313.x) (Accessed 2023-06-04).

Horner, John D., Steele, Cross J., Underwood, Christopher A.,
Lingamfelter, Daniel. "Age-Related Changes in Characteristics and Prey
Capture of Seasonal Cohorts of Sarracenia alata Pitchers." The American
Midland Naturalist, Vol. 167, No. 1, 2012,
[doi](https://doi.org/10.1674/0003-0031-167.1.13) (Accessed 2023-06-04).
